### 2021-08-24 paper/localpsf.tex
edits to regular grid transfer section based on discussionn yesterday with Noemi. Add further dictussion about why transferring to regular grid reduces costs

% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={Fast matrix-free approximation of smoothly varying blur operators, with application to Hessians in PDE-constrained inverse problems with highly informative data},
  pdfauthor={N. Alger, N. Petra, and O. Ghattas}
}
\fi

@@

To convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), we will need to evaluate $\varphi_{x_i}(z)$ for a large number of points $z$. Per \eqref{eq:varphi_eval}, evaluating $\varphi_{x_i}(z)$ requires evaluating $\eta_b(z+x_i)$. In computations, $\eta^b$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh. Evaluating $\eta_b(z+x_i)$ therefore requires finding which mesh cell that $z+x_i$ is in, then evaluating several polynomial functions to evaluate the finite element function on that cell. Although this procedure is asymptotically scalable, it is slow in practice \cite{FEMEVALSLOW}.

To speed up evaluation of $\varphi_{x_i}$, we interpolate $\varphi_{x_i}$ onto a regular rectilinear grid on a coordinate axis aligned box that just barely contains the ellipsoid $E_{x_i}$. We transfer the function $\varphi_{x_i}$ from the irregular mesh to the regular grid by evaluating $\varphi_{x_i}$ at all gridpoints in the regular grid, using the slow process described in the previous paragraph. This is an upfront cost which is done ``offline.'' After $\varphi_{x_i}$ has been transferred to the regular grid, we use fast regular grid interpolation to perform all subsequent ``online'' evaluations of $\varphi_{x_i}$. This yields a substantial overall speedup because the number of gridpoints in the regular grid for $\varphi_{x_i}$ is typically much smaller than the number of evaluations of $\varphi_{x_i}$ required to convert $\AopPc$ to H-matrix format. DISCUSS MORE ABOUT COMPUTATIONAL COSTS/SAVINGS
%Having $\varphi_{x_i}$ stored on a regular grid also simplifies the implementation of the method that we use to deal with boundary artifacts, which will be described in Section \ref{sec:boundary_extension}.

>>>>

% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={Fast matrix-free approximation of smoothly varying blur operators, with application to Hessians in PDE-constrained inverse problems with highly informative data},
  pdfauthor={N. Alger, N. Petra, T. Hartland, and O. Ghattas}
}
\fi

@@

To convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), we will need to evaluate $\varphi_{x_i}(z)$ for a large number of points $z$. Per \eqref{eq:varphi_eval}, evaluating $\varphi_{x_i}(z)$ requires evaluating $\eta_b(z+x_i)$. In computations, $\eta^b$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh. Evaluating $\eta^b(z+x_i)$ therefore requires finding which mesh cell that $z+x_i$ is in, then evaluating several polynomial functions to evaluate the finite element function on that cell. Although this procedure is asymptotically scalable, it is slow in practice \cite{FEMEVALSLOW}.

To speed up evaluation of $\varphi_{x_i}$, we interpolate $\varphi_{x_i}$ onto a regular rectilinear grid on a coordinate axis aligned box that just barely contains the ellipsoid $E_{x_i}$. We transfer the function $\varphi_{x_i}$ from the irregular mesh to the regular grid by evaluating $\varphi_{x_i}$ at all gridpoints in the regular grid, using the slow process described in the previous paragraph. This is an upfront cost which is done ``offline.'' After $\varphi_{x_i}$ has been transferred to the regular grid, we use fast regular grid interpolation to perform all subsequent ``online'' evaluations of $\varphi_{x_i}$. This yields a substantial overall speedup because the number of finite element function evaluations required to transfer the functions $\varphi_{x_i}$ to the regular grid is $O(N)$, while the number of kernel evaluations required to convert $\AopPc$ to H-matrix format is $O(k N \log N)$. Here $N$ is the number of finite element degrees of freedom, and $k$ is the H-matrix rank. Furthermore, transferring $\varphi_{x_i}$ to a regular grid requires evaluating $\varphi_{x_i}$ at a predetermined and highly structured collection of points. This structure allows for faster determination of which mesh cells that points are in, and allows one to evaluate the finite element function at all regular gridpoints in parallel. In contrast, constructing the H-matrix requires evaluating $\varphi_{x_i}$ at an unpredictable sequence of scattered of points, which is an expensive task for finite element function evaluation, but a cheap task for regular grid interpolation. 
Having $\varphi_{x_i}$ stored on a regular grid also simplifies the implementation of the method that we use to deal with boundary artifacts, which will be described in Section \ref{sec:boundary_extension}.
